---
title: "R for Data Science notes and exercises"
output:
  html_document:
    df_print: paged
---

R Markdown Notebook to follow along the "R for Data Science" book by Hadley Wickham and Garrett Grolemund. Available from https://r4ds.had.co.nz/.

```{r}
library(tidyverse)

# Get a feel for the columns in the data.frame
str(mpg)

# First visualization, trying to understand correlation between mileage and displacement
ggplot(data = mpg, aes(x = displ, y = hwy)) + 
  geom_point()
```

### 3.2.4 Exercises
#### 1. Run ggplot(data = mpg). What do you see?

```{r}
ggplot(data = mpg)
```

#### 2. How many rows are in mpg? How many columns?

```{r}
nrow(mpg)
ncol(mpg)
```

#### 3. What does the drv variable describe? Read the help for ?mpg to find out.

```{r}
table(mpg$drv)
ggplot(data = mpg, aes(x = drv)) + geom_bar()
# Looks like 4WD, RWD, and FWD
ggplot(data = mpg, aes(x = drv, y = hwy)) + 
  geom_point()
# Better mileage in FWD (which there was a better way to compute these)
summary(mpg[which(mpg$drv == '4'),]$hwy)   # 19.17
summary(mpg[which(mpg$drv == 'f'),]$hwy) # 28.16
summary(mpg[which(mpg$drv == 'r'),]$hwy) # 21
# Found better way to compute in "Análise e Exploração de Dados com R" by Miguel Rocha and Pedro G. Ferreira
tapply(mpg$hwy, mpg$drv, mean)
# Found a better way to do it
?mpg # "the type of drive train, where f = front-wheel drive, r = rear wheel drive, 4 = 4wd"
```

#### 4. Make a scatterplot of hwy vs cyl.

```{r}
ggplot(data = mpg) +
  geom_point(aes(x = hwy, y = cyl))

# Looks much better
ggplot(data = mpg) +
  geom_point(aes(x = cyl, y = hwy))
```

#### 5. What happens if you make a scatterplot of class vs drv? Why is the plot not useful?

```{r}
ggplot(data = mpg) +
  geom_point(aes(x = class, y = drv))
ggplot(data = mpg) +
  geom_point(aes(y = class, x = drv))
# I consider this useful, you can how certain classes of cars have more different types of drive (there all kinds of subcompacts), compared to others (minivan only FWD) 
```


